Interval-Valued Quantum Probabilities: Contextuality and the Born Rule

### 2017-05-16 prop_letter.tex
major reorganization

Our first observation is that essential objectives of our program to
achieve a computable theory of quantum mechanics and finite-resource
measurement, attempted previously with computable number
systems~\cite{usat,geometry2013apsrev4,DQT2014}, may be achievable byextending classical interval-valued probability
measures~\cite{JamisonLodwick2004} to the quantum domain.  We begin by
axiomatizing a quantum interval-valued probability measure (QIVPM)
framework exploiting the fact that a density matrix approach can be
considered as equivalent to a description of quantum states using
probability
measures~\cite{10.2307/2308516,Varadarajan2008}. %Mackey=10.2307
One result is that the expectation value of an observable with respect
to a QIVPM can be defined to be consistent with the constraints on the
expectation value of the same observable expected from continuous
quantum mechanics.  We then examine Gleason's
theorem~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum} and the
Kochen-Specker theorem in the context of QIVPM's, and conclude
that they correspond to the limit of an infinite number of precise
intervals, and the set of crypto-deterministic intervals (the set
$\{\imposs,\necess\}$), respectively.

@@

As an example, consider a three-dimensional Hilbert space with
orthonormal basis $\{\ket{0},\ket{1},\ket{2}\}$ and an observable $Q$
with spectral decomposition $\proj{0} + 2\,\proj{1} +
3\,\proj{2}$. Two fragments of valid probability measures $\mu_1$ and
$\mu_2$ that can associated with this space: $\mu_1(\proj{0}) = 0.5$,
$\mu_1(\proj{1}) = 0.5$, $\mu_1(\proj{2}) = 0$, and
$\mu_2(\proj{0}) = 0.5$, $\mu_2(\proj{1}) = 0$,
$\mu_2(\proj{2}) = 0.5$. By the Born rule, the first probability
measure corresponds to the quantum system being in the state
$(\ket{0}+\alpha\ket{1})/\sqrt{2}$ and the second corresponds to the
quantum system being in the state $(\ket{0}+\alpha\ket{2})/\sqrt{2}$,
where $|\alpha|=1$. The expectation value of the observable in the
first case is 1.5 and in the second it is 2.

The above axioms describe a mathematical idealization involving
infinitely precise measurements. In an actual experimental setup with
an ensemble of quantum states that are not exactly identical, and with
small imperfections and inaccuracies in measuring devices, an
experimenter might only be able to determine that the probability of
the event $\proj{0}$ is in the range $[0.49,0.51]$. The spread in this
range depends on the amount of resources (time, energy, money, etc)
that are devoted to the experiment. In the classical setting, this
``fuzziness'' is formalized by moving to \emph{interval-valued
  probability measures}.

Given a collection of intervals $\mathscr{I}$ with addition and scalar
multiplication defined as follows: 
\begin{subequations}\label{eq:interval-operations}
\begin{eqnarray}
 &  & [l_{1},r_{1}]+[l_{2},r_{2}]=[l_{1}+l_{2},r_{1}+r_{2}]\textrm{ and}\\
 &  & x[l,r]=\begin{cases}
[xl,xr] & \textrm{for }x\ge0\textrm{ ;}\\{}
[xr,xl] & \textrm{for }x\le0\textrm{ ,}
\end{cases}
\end{eqnarray}
\end{subequations}
a QIVPM $\bar{\mu}$ is an assignment of intervals to each event
obeying the following axioms:
$\bar{\mu}(\mathbb{0})=\imposs=\left[0,0\right]$,
$\bar{\mu}(\mathbb{1})=\necess=\left[1,1\right]$,
$\bar{\mu}\left(\mathbb{1}-P\right)=\left[1,1\right]-\bar{\mu}\left(P\right)$,
and for a set of mutually orthogonal projections
$\left\{ P_{i}\right\} _{i=1}^{N}$, we have:
\begin{equation}
\bar{\mu}\left(\sum_{i=1}^{N}P_{i}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(P_{i}\right)\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
The first three axioms are the direct counterpart of the corresponding
axioms for conventional probability measures. The last condition
allows for the interval associated with a composite event to be
sharper than the interval calculated from each individual
event. Intuitively, consider an event $P$ and its complement
$\mathbb{1}-P$. Looking at each separately we may discover as a result
of an experiment that $\bar{\mu}(P) = [0.2,0.3]$ and
$\bar{\mu}(\mathbb{1}-P)=[0.7,0.8]$. By adding the individual
intervals, we predict the probability of the composite event
$P+(\mathbb{1}-P)$ to be in the range
$[0.9,1.1]$. Eq.~\ref{eq:QuantumInterval-valuedProbability-Inclusion}
allows us to refine this to $[1,1]$ by recognizing that the composite
event is the identity projection.

@@

@@

Our first result is that a QIVPM cannot be crypto-deterministic. We
will use the proof of the Kochen-Specker
theorem~\cite{kochenspecker1967,peres1995quantum,Redhead1987-REDINA,Griffiths2003}
to show that it is, in general, impossible to just have the two
intervals \imposs\ and \necess: every QIVPM must be parameterized by a
collection of intervals that includes at least one ``fuzzy'' interval
and some events must be associated with these uncertain intervals.

@@

\amr{does the following rely on convexity?}

\newpage

Suppose there is a quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\mathrm{D}}$.
We can prove $\expval{\mathbf{O}}_{\bar{\mu}}$ is a singleton set~$\left\{ \lambda\right\} $,
where $\lambda$ is an eigenvalue of $\mathbf{O}$. If we define the
multiplication of two singleton sets~$\left\{ \lambda_{1}\right\} $
and $\left\{ \lambda_{2}\right\} $ to be $\left\{ \lambda_{1}\lambda_{2}\right\} $,
then we have 
\begin{equation}
\prod_{j=1}^{N}\expval{\mathbf{O}_{j}}_{\bar{\mu}}=\expval{\prod_{j=1}^{N}\mathbf{O}_{j}}_{\bar{\mu}}\label{eq:Kochen-Specker-Rules}
\end{equation}
for any sequence of commuting observables~$\left\{ \mathbf{O}_{j}\right\} _{j=1}^{N}$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Convexity}

@@

Since our measurement is fuzzy, generally there are more than one
state consistent with a particular quantum IVPM. For example, consider
three intervals $\left[0,0\right]$, $\left[1,1\right]$ and \emph{$\left[0,1\right]$},
where $\left[0,0\right]$ and $\left[1,1\right]$ are called \imposs~and
\necess~as before, and \emph{$\left[0,1\right]$} is called \unknown~because
it provides no information. Then,
\begin{equation}
\bar{\mu}_{1}\left(P\right)=\begin{cases}
\necess & \text{, if }P=\mathbb{1}\text{ ;}\\
\imposs & \text{, if }P=\mathbb{0}\text{ ;}\\
\unknown & \text{, otherwise,}
\end{cases}\label{eq:no-information-IVPM}
\end{equation}
is a quantum IVPM. This measure represents the beliefs of an experimenter
with no prior knowledge about the particular quantum system in question,
and any state is consistent with $\bar{\mu}_{1}$ on any projectors.

Not specific to quantum; fuziness also in classical setting

Quantum IVPMs become classical IVPMs~\cite{JamisonLodwick2004} whenthe space of events includes only mutually commuting sub-events, where
$\events'\subseteq\events$ is called a sub-event space if it satisfies
the following conditions.
\begin{itemize}
\item $\left\{ \mathbb{0},\mathbb{1}\right\} \subseteq\events'$.
\item For any projection~$P\in\events'$, we have $\mathbb{1}-P\in\events'$. 
\item For a set of mutually orthogonal projections~$\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\events'$,
we have $\sum_{i=1}^{N}P_{i}\in\events'$. 
\end{itemize}
For example, $\events_{0}=\left\{ \mathbb{0},\proj{0},\proj{1},\mathbb{1}\right\} $
is a commuting sub-event space. Restricting $\bar{\mu}_{1}$ on $\events_{0}$,
$\bar{\mu}_{1}|_{\events_{0}}$, gives a classical IVPM~$\bar{\mu}_{2}$:
\begin{equation}
\begin{aligned}\bar{\mu}_{2}(\mathbb{0}) & =\imposs & \bar{\mu}_{2}(\proj{0}) & =\unknown\\
\bar{\mu}_{2}(\mathbb{1}) & =\necess & \bar{\mu}_{2}(\proj{1}) & =\unknown
\end{aligned}
\end{equation}
If we identify the density matrices~$\proj{0}$ and $\proj{1}$ as
the head and the tail of a coin, respectively, then $\bar{\mu}_{2}$
represents that an experimenter knows nothing about a coin before
she tosses it. Moreover, for each $p\in\left[0,1\right]$, the mixed
state~$p\proj{0}+\left(1-p\right)\proj{1}$ represents a classical
coin with probability~$p$ to get head, and $p\proj{0}+\left(1-p\right)\proj{1}$
is consistent with $\bar{\mu}_{2}$ on $\events_{0}$ because any
classical coin should be consistent with the situation where an experimenter
has no prior knowledge.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fuzzy Observables}

@@

Formally, the expectation value of an observable with respect to an
IVPM should be the minimum and maximum expectation value of the same
observable with respect to states consistent with it. This is true
when we require our IVPM to be convex, and on a commuting sub-event
space induced by an observable.

\begin{definition}Given an observable~$\mathbf{O}$ with its spectrum
decomposition~(\ref{eq:spectrum-decomposition}). A set of projectors~$\events'\subseteq\events$
is called a sub-event space induced by $\mathbf{O}$ if it is the
minimal sub-event space containing $\left\{ P_{i}^{-}\right\} _{i=1}^{N^{-}}\cup\left\{ P_{i}^{+}\right\} _{i=1}^{N^{+}}$.\end{definition}

\begin{thm}Given a Hilbert space of $d\ge3$, for every quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
and any observable~$\mathbf{O}$ inducing a sub-event space~$\events'$,
if $\bar{\mu}$ is convex, then
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\left[\min_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho},\max_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho}\right]\textrm{ .}\label{eq:quantum-interval-expectation-core}
\end{equation}
Conversely, if Eq.~(\ref{eq:quantum-interval-expectation-core})
holds for every observable, then $\bar{\mu}$ is convex.\end{thm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fuzzy States}

>>>>

Our first observation is that essential objectives of our program to
achieve a computable theory of quantum mechanics and finite-resource
measurement, attempted previously with computable number
systems~\cite{usat,geometry2013apsrev4,DQT2014}, may be achievable by
extending classical interval-valued probability
measures~\cite{JamisonLodwick2004} to the quantum domain.  We begin by
axiomatizing a quantum interval-valued probability measure (QIVPM)
framework exploiting the fact that a density matrix approach can be
considered as equivalent to a description of quantum states using
probability
measures~\cite{10.2307/2308516,Varadarajan2008}. %Mackey=10.2307
One result is that the expectation value of an observable with respect
to a QIVPM can be defined to be consistent with the constraints on the
expectation value of the same observable expected from continuous
quantum mechanics.  We then examine Gleason's
theorem~\cite{gleason1957,Redhead1987-REDINA,peres1995quantum} and the
Kochen-Specker theorem in the context of QIVPM's, and conclude that
they correspond to the limit of an infinite number of precise
intervals, and the set of crypto-deterministic intervals (the set
$\{\imposs,\necess\}$), respectively.

@@

As an example, consider a three-dimensional Hilbert space with
orthonormal basis $\{\ket{0},\ket{1},\ket{2}\}$ and an observable $Q$
with spectral decomposition $\proj{0} + 2\,\proj{1} +
3\,\proj{2}$. Two fragments of valid probability measures $\mu_1$ and
$\mu_2$ that can associated with this space are:
$\mu_1(\proj{0}) = 0.5$, $\mu_1(\proj{1}) = 0.5$,
$\mu_1(\proj{2}) = 0$, and $\mu_2(\proj{0}) = 0.5$,
$\mu_2(\proj{1}) = 0$, $\mu_2(\proj{2}) = 0.5$. By the Born rule, the
first probability measure corresponds to the quantum system being in a
state $(\ket{0}+\alpha\ket{1})/\sqrt{2}$ and the second corresponds to
the quantum system being in the state
$(\ket{0}+\alpha\ket{2})/\sqrt{2}$, where in both cases
$|\alpha|=1$. The expectation value of the observable in the first
case is 1.5 and in the second it is 2.

The above axioms describe a mathematical idealization involving
infinitely precise measurements. In an actual experimental setup with
an ensemble of quantum states that are not exactly identical, and with
small imperfections and inaccuracies in measuring devices, an
experimenter might only be able to determine that the probability of
the event $\proj{0}$ is in the range $[0.49,0.51]$ instead of
precisely 0.5. The spread in this range depends on the amount of
resources (time, energy, money, etc) that are devoted to the
experiment. In the classical setting, this ``fuzziness'' is formalized
by moving to \emph{interval-valued probability measures}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Intervals of Uncertainty}

We will start by reviewing IVPMs in the classical
setting~\cite{JamisonLodwick2004} and then propose our quantum
framework. In the classical setting, there are several proposals for
``imprecise probabilities.'' Although these proposals differ in some
details, they all share the fact that the probability $\mu(E)$ of an
event $E$ is generalized from a single real number to an interval of
certainty $[\ell,r]$ where~$\ell$ intuitively corresponds to the
strength of evidence for the event and $r$ corresponds to the strength
of evidence against the event. The collection of intervals used in the
definition of a probability measure must satisfy certain properties
which we explain next.

First for each interval $[\ell,r]$ we have the natural constraint
$0 \leq \ell \leq r \leq 1$ that guarantees that every element of the
interval can be interpreted as a conventional probability. We also
include $\imposs=[0,0]$ and $\necess=[1,1]$ as limiting intervals to
refer to the probability of impossible events (e.g., the null event)
and certain events (e.g., the event covering the entire space). For
each interval $[\ell,r]$, we also need the interval $[1-r,1-\ell]$ so
that if one interval refers to the probability of an event, the other
refers to the probability of the complement event. For example, if we
discover as a result of an experiment that $\mu(E) = [0.2,0.3]$ for
some event $E$, we may conclude that $\mu(\overline{E}) = [0.7,0.8]$
for the complement event $\overline{E}$. In addition to these simple
conditions, there are some subtle conditions on how intervals are
combined, which we discuss next.

Let $E_1$ and $E_2$ be to disjoint events with probabilities
$\mu(E_1)=[\ell_1,r_1]$ and $\mu(E_2)=[\ell_2,r_2]$. A first attempt
at calculating the probability of the combined event that $E_1$ or
$E_2$ occurs might be $\mu(E_1+E_2) = [\ell_1+\ell_2,r_1+r_2]$. In
some cases, this is indeed a sensible definition. For example, if
$\mu(E_1)=[0.1,0.2]$ and $\mu(E_2)=[0.3,0.4]$ we get
$\mu(E_1+E_2) = [0.4,0.6]$. But consider an event $E$ such that
$\mu(E)=[0.2,0.3]$ and hence $\mu(\overline{E})=[0.7,0.8]$. The two
events are disjoint; the previous na\"\i ve way of combining the
intervals would give $\mu(E+\overline(E))=[0.9,1.1]$ which is not a
valid probability interval. Moreover the event $E+\overline{E}$ is the
identity event covering the entire space; its probability interval
should be $[1,1]$ which is sharper than $[0.9,1.1]$. The problem is
that the two intervals are correlated: there is more information in
the combined event than in each event separately. In our example, even
though the true probability of $\mu(E)$ can be in anywhere in the
range $[0.2,0.3]$ and the true probability of $\mu(\overline{E})$ can
be in anywhere in the range $[0.7,0.8]$, the values are not
independent. Any value of $\mu(E) \leq 0.25$ will force
$\mu(\overline{E})\geq 0.75$. To account for such subtleties, the
axioms of interval-valued probability do not use a strict equality for
the combination of disjoint events. The correct constraint regarding
the probability of $E_1+E_2$ when $E_1$ and $E_2$ are disjoint is
$\mu(E_1+E_2) \subseteq [\ell_1+\ell_2,r_1+r_2]$ which, in combination
with the requirement for the identity event would force
$\mu(E+\overline{E})$ to always be $[1,1]$ irrespective of the
particular assignments of $\mu(E)$ or $\mu(\overline{E})$. When
combining non-disjoint events, there is a further subtlety whose
resolution will give the most general and final condition as
follows. For general, non-necessarily disjoint events $E_1$ and $E_2$,
we have:
\begin{equation}
\mu(E_1+E_2) + \mu(E_1E_2) \subseteq \mu(E_1) + \mu(E_2)
\label{eq:classicalconvex}
\end{equation}
which is a generalization of the classical inclusion-exclusion
principle that uses $\subseteq$ instead of $=$ for the same reason as
before. The new condition reduces to the previously motivated
condition when the events are disjoint, i.e., when $\mu(E_1E_2) = 0$.

We now have the necessary ingredients to define QIVPMs as a
generalization of the above observations. More precisely the quantum
definition will reduce to the classical one when the space of quantum
events is restricted to mutually commuting sub-events.

\begin{definition}[QIVPM]
  Given a collection of intervals $\mathscr{I}$ including
  $\imposs=[0,0]$ and $\necess=[1,1]$ and with addition and scalar
  multiplication defined as follows:
  \begin{subequations}\label{eq:interval-operations}
  \begin{eqnarray}
   &  & [l_{1},r_{1}]+[l_{2},r_{2}]=[l_{1}+l_{2},r_{1}+r_{2}]\textrm{ and}\\
   &  & x[l,r]=\begin{cases}
  [xl,xr] & \textrm{for }x\ge0\textrm{ ;}\\{}
  [xr,xl] & \textrm{for }x\le0\textrm{ ,}
  \end{cases}
  \end{eqnarray}
  \end{subequations}
  a QIVPM $\bar{\mu}$ is an assignment of intervals to each event
  (projection operator) obeying the following axioms:
  $\bar{\mu}(\mathbb{0})=\imposs=\left[0,0\right]$,
  $\bar{\mu}(\mathbb{1})=\necess=\left[1,1\right]$,
  $\bar{\mu}\left(\mathbb{1}-P\right)=\left[1,1\right]-\bar{\mu}\left(P\right)$,
  and for each pair of \emph{commuting} projections $P_0$ and $P_1$,
  we have:
\begin{equation}
\bar{\mu}\left(P_{0}+P_{1}-P_{0}P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)
\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
\end{definition}
The first three axioms are the direct counterpart of the corresponding
axioms for conventional probability measures. With the understanding
that the union of classical sets $E_1+E_2$ is replaced by
$P_0+P_1-P_0P_1$ in the case of projection operators, the last
condition, Eq~\ref{eq:QuantumInterval-valuedProbability-Inclusion},
is a direct counterpart of Eq.~\ref{eq:classicalconvex}.

Quantum IVPMs become classical IVPMs when the space of events includes
only mutually commuting sub-events, where $\events'\subseteq\events$
is called a sub-event space if it satisfies the following conditions.
\begin{itemize}
\item $\left\{ \mathbb{0},\mathbb{1}\right\} \subseteq\events'$.
\item For any projection~$P\in\events'$, we have $\mathbb{1}-P\in\events'$. 
\item For a set of mutually orthogonal projections~$\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\events'$,
we have $\sum_{i=1}^{N}P_{i}\in\events'$. 
\end{itemize}

For example, assuming the three intervals $\imposs$, $\necess$, and
$\unknown=\left[0,1\right]$, then:
\begin{equation}
\bar{\mu}_{1}\left(P\right)=\begin{cases}
\necess & \text{, if }P=\mathbb{1}\text{ ;}\\
\imposs & \text{, if }P=\mathbb{0}\text{ ;}\\
\unknown & \text{, otherwise,}
\end{cases}\label{eq:no-information-IVPM}
\end{equation}
is a quantum IVPM. This measure represents the beliefs of an
experimenter with no prior knowledge about the particular quantum
system in question, and any state is consistent with $\bar{\mu}_{1}$
on any projectors. We have that
$\events_{0}=\left\{ \mathbb{0},\proj{0},\proj{1},\mathbb{1}\right\} $
is a commuting sub-event space. Restricting $\bar{\mu}_{1}$ on
$\events_{0}$, $\bar{\mu}_{1}|_{\events_{0}}$, gives a classical
IVPM~$\bar{\mu}_{2}$:
\begin{equation}
\begin{aligned}\bar{\mu}_{2}(\mathbb{0}) & =\imposs & \bar{\mu}_{2}(\proj{0}) & =\unknown\\
\bar{\mu}_{2}(\mathbb{1}) & =\necess & \bar{\mu}_{2}(\proj{1}) & =\unknown
\end{aligned}
\end{equation}
If we identify the density matrices~$\proj{0}$ and $\proj{1}$ as
the head and the tail of a coin, respectively, then $\bar{\mu}_{2}$
represents that an experimenter knows nothing about a coin before
she tosses it. Moreover, for each $p\in\left[0,1\right]$, the mixed
state~$p\proj{0}+\left(1-p\right)\proj{1}$ represents a classical
coin with probability~$p$ to get head, and $p\proj{0}+\left(1-p\right)\proj{1}$
is consistent with $\bar{\mu}_{2}$ on $\events_{0}$ because any
classical coin should be consistent with the situation where an experimenter
has no prior knowledge.

@@

Our first result regarding QIVPM is that a QIVPM cannot be
crypto-deterministic in the sense of Peres. We will use the proof of
the Kochen-Specker
theorem~\cite{kochenspecker1967,peres1995quantum,Redhead1987-REDINA,Griffiths2003}
to show that it is, in general, impossible to just have the two
intervals \imposs\ and \necess: every QIVPM must be parameterized by a
collection of intervals that includes at least one additional
``fuzzy'' interval and some events must be associated with such
uncertain intervals. The significance of this observation is that the
uncertainty in the intervals is, in contrast to the classical case,
\emph{a fundamental part of quantum theory}. There is in fact two
sources of imprecision: one due to the finite measurement resources
and one due to complementarity which requires every QIVPM to be truly
probabilistic.

@@

The argument refers to the expectation value of these
observables. In the case of QIVPM this is defined as follows.

\begin{definition} Given an observable~$\mathbf{O}$ with spectral
  decomposition
  $\mathbf{O}=\sum_{s\in\left\{ +,-\right\}
  }\sum_{i=1}^{N^{s}}\lambda_{i}^{s}P_{i}^{s}$, a set of
  projectors~$\events'\subseteq\events$ is called a sub-event space
  induced by $\mathbf{O}$ if it is the minimal sub-event space
  containing
  $\left\{ P_{i}^{-}\right\} _{i=1}^{N^{-}}\cup\left\{
    P_{i}^{+}\right\} _{i=1}^{N^{+}}$.
\end{definition}

\begin{thm} Given a Hilbert space of $d\ge3$, for every
  QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$ and any
  observable~$\mathbf{O}$ inducing a sub-event space~$\events'$, 
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\left[\min_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho},\max_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho}\right]\textrm{ .}\label{eq:quantum-interval-expectation-core}
\end{equation}
\end{thm}

The expectation value is intuitively between two possible outcomes
which lie between the minimum and maximum bounds of the given
probability intervals. If the QIVPM is crypto-deterministic, this
interval collapses to a singleton $\left\{ \lambda\right\} $, where
$\lambda$ is an eigenvalue of $\mathbf{O}$. If we define the
multiplication of two singleton sets~$\left\{ \lambda_{1}\right\} $
and $\left\{ \lambda_{2}\right\} $ to be
$\left\{ \lambda_{1}\lambda_{2}\right\} $, then we have
\begin{equation}
\prod_{j=1}^{N}\expval{\mathbf{O}_{j}}_{\bar{\mu}}=\expval{\prod_{j=1}^{N}\mathbf{O}_{j}}_{\bar{\mu}}\label{eq:Kochen-Specker-Rules}
\end{equation}
for any sequence of commuting observables~$\left\{ \mathbf{O}_{j}\right\} _{j=1}^{N}$.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage  
\section{Rest of Paper}

Consider again
Eq.~\ref{eq:QuantumInterval-valuedProbability-Inclusion}. We will
generalize it as follows for all commuting $P_{0},P_{1}\in\events$, we
have
$\bar{\mu}\left(P_{0}+P_{1}-P_{0}P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)$. Note
that if $P_0$ and $P_1$ are \emph{orthogonal}, i.e., if $P_0P_1=0$ the
new equation reduces to
Eq.~\ref{eq:QuantumInterval-valuedProbability-Inclusion}.

\begin{verbatim}
Inclusion exclusion principle 
 pr(e1 union e2) + pr (e1 intersect e2) = 
  pr(e1) + pr(e2)

in classical Real number prob. LHS = X; RHS = X

---

in classical IVPM; we only get LHS <= X; RHS <= X
but we have no connection between LHS and RHS

convexity imposes that LHS >= RHS

---

our condition in the quantum case is the same; 
our left(LHS) is >= left(RHS); the right is dual, 
which makes LHS interval subset RHS interval.

--

replace set union by projection sum = p0 + p1 - p0p1
replace = by subseteq

and we get convexity definition

--

Now we need a simple classical example that shows 
why we need convexity or at least what it means

must be dimension 4 because everything of dim 3 
is convex in classical case

Consider omega = { 0, 1, 2, 3 }
intervals = imposssible, certain, 
            unlikely = [0,0.5], likely [0.5,1]

mu(empty) = imposs
mu(omega) = certain
mu({j}) = unlikely
mu(Omega \ {j}) = likely
mu({0,1}) = mu({0,2}) = mu({1,2}) = unlikely
mu({0,3}) = mu({1,3}) = mu({2,3}) = likely

EQ A:
mu({0}) + mu({0,1,2}) = [0.5,1.5]
mu({0,1}) + mu({0,2}) = [0,1]

AND [0.5,1.5] is not subseteq of [0,1]

Note that there is a core p(0) = p(1) = p(2) = 0.2, 
p(3) = 0.4, and under that core both sides of EQ A 
have prob. 0.8. So one way to make it convex is 
to use very precise intervals [0.2,0.2] etc.

Another way to make this convex is by choosing 
just one additional interval unknown [0,1] 
instead of unlikely and likely.

Intuition: the more you group the more 
precise you get

\end{verbatim}

@@

@@

Intuitively, the expectation value of an observable with respect to a
QIVPM is bounded by the minimum and maximum expectation value of the
same observable with respect to states consistent with it. This is
true when we require our IVPM to be convex, and on a commuting
sub-event space induced by an observable.
